Minimizing post shock forecasting error using disparate information

### 2020-02-26 eck2019syntheticprediction.tex
updated synthetic control notes that borrow from Soheils direction
updated MS

\newcommand{\w}{\textbf{w}}
\newcommand{\x}{\textbf{x}}
\newcommand{\X}{\textbf{X}}
\newcommand{\Y}{\textbf{Y}}
\newcommand{\Hist}{\mathcal{H}}

@@

@@

  - get a real dataset. \\

@@

We will suppose that an analyst has time series data ($y_{i,t}$,$\x_{i,t}$), 
$t = 1$, $\ldots$, $T_i$, $i = 1$, $\ldots$, $n+1$, where $y_{i,t}$ is a 
scalar response and $\x_{i,t}$ is a vector of covariates that revealed prior 
to the observation of $y_{1,t}$.  Suppose that the analyst is interested in 
forecasting $y_{1,t}$, the first time series in the collection.
%$y_1, y_2, \ldots$.  
%Given each time point $t \geq 1$, let $\x_{1,t}$ be a vector of covariates 
%revealed prior to the observation of $y_{1,t}$.  
To gauge the performance of a procedure that produces forecasts 
$\{\hat y_{1,t}, t= 1,2,\ldots\}$ given time horizon $T_1$, we consider the 
average forecast risk
$$
  R_T = \frac{1}{T}\sum_{t=1}^T\E(\hat y_{1,t} - y_{1,t})^2
$$
in our analyses. In this article, we consider a similar dynamic panel data 
model with autoregressive structure to that in \citet{blundell1998initial}. 
%without past time period covariate 
%information and with a single present time covariate.  
Our dynamic panel model includes an additional shock effect whose presence 
or absence is given by the binary variable $D_{i,t}$.  Our dynamic panel 
model is 
\begin{equation} \label{DPM}
  y_{i,t} = \eta_i + \alpha_iD_{i,t} + \phi_i y_{i,t-1} + \theta_i' \x_{i,t} + 
    \beta_i' \x_{i,t-1} + \varepsilon_{i,t},  
\end{equation}
where $t = 1,\ldots,T_i$, $i = 1,\ldots, n+1$, and $D_{i,t} = 1(t > T_i^*)$, 
$T_i^* < T_i$.  We will consider the following random effects structure:  
%\begin{align*}
%  \eta_i &\overset{iid}{\sim} N(0, \sigma_\eta^2), 
%    \qquad i = 1,\ldots n+1, \\
%  \alpha_i &\overset{iid}{\sim} N(\mu_\alpha, \sigma_\alpha^2), 
%    \qquad i = 1,\ldots n+1, \\
%  \phi_i &\overset{iid}{\sim} U(-1,1), 
%    \qquad i = 1,\ldots n+1, \\
%  \theta_i &\overset{iid}{\sim} N(0, \Sigma_\theta), 
%    \qquad i = 1,\ldots n+1, \\
%  \beta_i &\overset{iid}{\sim} N(0, \Sigma_\beta), 
%    \qquad i = 1,\ldots n+1, \\
%  \varepsilon_{i,t} &\overset{iid}{\sim} N(0, \sigma^2), 
%    \qquad i = 1,\ldots n+1;\; t = 1, \ldots T, \\
%  \eta &\indep \alpha \indep \phi \indep \theta \indep \varepsilon;
%\end{align*}
\begin{align*}
  \eta_i &\overset{iid}{\sim} \eta,\; \text{where}\; \E(\eta) = 0,\; \Var(\eta) = \sigma_\eta^2, 
    \qquad i = 1,\ldots n+1, \\
  \alpha_i &\overset{iid}{\sim} \alpha,\; \text{where}\; \E(\alpha) = \mu_\alpha,\; 
    \Var(\alpha) = \sigma_\alpha^2, \qquad i = 1,\ldots n+1, \\
  \phi_i &\overset{iid}{\sim} \phi,\; \text{where}\; |\phi| < 1, 
    \qquad i = 1,\ldots n+1, \\
  \theta_i &\overset{iid}{\sim} \theta,\; \text{where}\; \E(\theta) = 0,\; 
    \Var(\theta) = \Sigma_\theta^2, \qquad i = 1,\ldots n+1, \\
  \beta_i &\overset{iid}{\sim} \beta,\; \text{where}\; \E(\beta) = 0,\; 
    \Var(\beta) = \Sigma_\beta^2, \qquad i = 1,\ldots n+1, \\    
  \varepsilon_{i,t} &\overset{iid}{\sim} N(0, \sigma^2), 
    \qquad t = 1, \ldots T_i,\; i = 1,\ldots n+1; \\
  \eta &\indep \alpha \indep \phi \indep \theta \indep \varepsilon.
\end{align*}

@@

When we estimate all of the parameters via OLS we can write 
$\Var(\hat{\alpha}_{\text{adj}})$ as
\begin{equation}
\begin{split}
  &\Var(\hat{\alpha}_{\text{adj}}) = \E\left\{\Var(\hat{\alpha}_{\text{adj}}|\Hist)\right\} 
    + \Var\left\{\E(\hat{\alpha}_{\text{adj}}|\Hist)\right\} \\
    %= \E\left\{\Var(\hat{\alpha}_{\text{adj}}|\Hist)\right\} 
    %+ \frac{\sigma_{\alpha}^2}{n} \\
  &\qquad= \frac{1}{n^2} \sum_{i=2}^{n+1} \E\left\{\Var(\hat{\alpha}_i|\Hist)\right\} 
    + \frac{\sigma_{\alpha}^2}{n} \\
  &\qquad= \frac{1}{n^2(T-1)}\sum_{i=2}^{n+1} 
    \E\left\{\frac{s_{e,i}^2}{(1-R^2_i)s_{\alpha_i}^2}\right\} 
      + \frac{\sigma_{\alpha}^2}{n},
\label{reg}
\end{split}
\end{equation}
where $s_{e,i}^2$ is the estimated variance of model $i$, $R^2_{\alpha_i}$ is 
the multiple $R^2$ obtained from regressing the shock indicator $D$ on the 
other regressors, and $s_{\alpha_i}^2$ is... .

Propostion~\ref{prop:R1R2} states that Forecast 2 has lower forecast risk 
than Forecast 1 when  
$
  \Var(\hat{\alpha}_{\text{adj}}) 
    %- 2\E\left(\hat\eta_1 + \hat\phi_1 y_{1,T^*} + \hat\theta_1'\x_{1,T^*+1} 
    %+ \hat\beta_1'\x_{1,T^*} - \eta_1 - \phi_1 y_{1,T^*} - \theta_1'\x_{1,T^*+1}
    %- \beta_1'\x_{1,T^*}\right)\mu_{\alpha} 
    < \mu_{\alpha}^2.
$
Combining this result with \eqref{reg}, we see that Forecast 2 has lower 
forecast risk than Forecast 1 when 
$$
  \frac{1}{n^2(T-1)}\sum_{i=2}^{n+1} \E\left\{\frac{s_{e,j}^2}{(1-R^2_i)s_{\alpha_i}^2}\right\} 
    + \frac{\sigma_{\alpha}^2}{n} < \mu_{\alpha}^2.
$$
Forecast 2 is preferable to Forecast 1 asymptotically in both $T$ and $n$ 
whenever $\mu_{\alpha} \neq 0$. In finite samples, Forecast 2 is preferable to 
Forecast 1 when the $\mu_{\alpha}$ is large relative to its variability and 
overall regression variability.

>>>>

\newcommand{\R}{\mathbb{R}}
\newcommand{\w}{\textbf{w}}
\newcommand{\x}{\textbf{x}}
\newcommand{\X}{\textbf{X}}
\newcommand{\Y}{\textbf{Y}}
\newcommand{\Hist}{\mathcal{H}}

@@

  - more general panel model where covariates may be different. \\

@@

  - get a real dataset. Air conditiong units in hot humid summer vs salt sales 
  	in very icy/snoy winters and portable generators for hurricane/fires/tornados  \\

@@

We will suppose that an analyst has time series data ($y_{i,t}$,$\x_{i,t}$), 
$t = 1$, $\ldots$, $T_i$, $i = 1$, $\ldots$, $n+1$, where $y_{i,t}$ is a 
scalar response and $\x_{i,t}$ is a vector of covariates that revealed prior 
to the observation of $y_{1,t}$.  Suppose that the analyst is interested in 
forecasting $y_{1,t}$, the first time series in the collection.
%$y_1, y_2, \ldots$.  
%Given each time point $t \geq 1$, let $\x_{1,t}$ be a vector of covariates 
%revealed prior to the observation of $y_{1,t}$.  
To gauge the performance of a procedure that produces forecasts 
$\{\hat y_{1,t}, t= 1,2,\ldots\}$ given time horizon $T_1$, we consider the 
average forecast risk
$$
  R_T = \frac{1}{T}\sum_{t=1}^T\E(\hat y_{1,t} - y_{1,t})^2
$$
in our analyses. In this article, we consider a similar dynamic panel data 
model with autoregressive structure to that in \citet{blundell1998initial}. 
%without past time period covariate 
%information and with a single present time covariate.  
Our dynamic panel model includes an additional shock effect whose presence 
or absence is given by the binary variable $D_{i,t}$.  Our dynamic panel 
model is 
\begin{equation} \label{DPM}
  y_{i,t} = \eta_i + \alpha_iD_{i,t} + \phi_i y_{i,t-1} + \theta_i' \x_{i,t} + 
    \beta_i' \x_{i,t-1} + \varepsilon_{i,t},  
\end{equation}
$t = 1,\ldots,T_i$ and $i = 1,\ldots, n+1$, where $D_{i,t} = 1(t > T_i^*)$, 
$T_i^* < T_i$ and $\x_{i,t} \in \R^{p}$, $p \geq 1$.  We will consider 
the following random effects structure:  
%\begin{align*}
%  \eta_i &\overset{iid}{\sim} N(0, \sigma_\eta^2), 
%    \qquad i = 1,\ldots n+1, \\
%  \alpha_i &\overset{iid}{\sim} N(\mu_\alpha, \sigma_\alpha^2), 
%    \qquad i = 1,\ldots n+1, \\
%  \phi_i &\overset{iid}{\sim} U(-1,1), 
%    \qquad i = 1,\ldots n+1, \\
%  \theta_i &\overset{iid}{\sim} N(0, \Sigma_\theta), 
%    \qquad i = 1,\ldots n+1, \\
%  \beta_i &\overset{iid}{\sim} N(0, \Sigma_\beta), 
%    \qquad i = 1,\ldots n+1, \\
%  \varepsilon_{i,t} &\overset{iid}{\sim} N(0, \sigma^2), 
%    \qquad i = 1,\ldots n+1;\; t = 1, \ldots T, \\
%  \eta &\indep \alpha \indep \phi \indep \theta \indep \varepsilon;
%\end{align*}
\begin{align*}
  \eta_i &\overset{iid}{\sim} \eta,\; \text{where}\; \E(\eta) = 0,\; \Var(\eta) = \sigma_\eta^2, 
    \qquad i = 1,\ldots n+1, \\
  \alpha_i &\overset{iid}{\sim} \alpha,\; \text{where}\; \E(\alpha) = \mu_\alpha,\; 
    \Var(\alpha) = \sigma_\alpha^2, \qquad i = 1,\ldots n+1, \\
  \phi_i &\overset{iid}{\sim} \phi,\; \text{where}\; |\phi| < 1, 
    \qquad i = 1,\ldots n+1, \\
  \theta_i &\overset{iid}{\sim} \theta,\; \text{where}\; \E(\theta) = 0,\; 
    \Var(\theta) = \Sigma_\theta^2, \qquad i = 1,\ldots n+1, \\
  \beta_i &\overset{iid}{\sim} \beta,\; \text{where}\; \E(\beta) = 0,\; 
    \Var(\beta) = \Sigma_\beta^2, \qquad i = 1,\ldots n+1, \\    
  \varepsilon_{i,t} &\overset{iid}{\sim} N(0, \sigma^2), 
    \qquad t = 1, \ldots T_i,\; i = 1,\ldots n+1; \\
  \eta &\indep \alpha \indep \phi \indep \theta \indep \varepsilon.
\end{align*}

@@

When we estimate all of the parameters via OLS we can write 
$\Var(\hat{\alpha}_{\text{adj}})$ as
\begin{equation}
\begin{split}
  &\Var(\hat{\alpha}_{\text{adj}}) = \E\left\{\Var(\hat{\alpha}_{\text{adj}}|\Hist)\right\} 
    + \Var\left\{\E(\hat{\alpha}_{\text{adj}}|\Hist)\right\} \\
    %= \E\left\{\Var(\hat{\alpha}_{\text{adj}}|\Hist)\right\} 
    %+ \frac{\sigma_{\alpha}^2}{n} \\
  &\qquad= \frac{1}{n^2} \sum_{i=2}^{n+1} \E\left\{\Var(\hat{\alpha}_i|\Hist)\right\} 
    + \frac{\sigma_{\alpha}^2}{n}. %\\
%  &\qquad= \frac{1}{n^2(T-1)}\sum_{i=2}^{n+1} 
%    \E\left\{\frac{s_{e,i}^2}{(1-R^2_i)s_{\alpha_i}^2}\right\} 
%      + \frac{\sigma_{\alpha}^2}{n},
\label{reg}
\end{split}
\end{equation}
%where $s_{e,i}^2$ is the estimated variance of model $i$, $R^2_{\alpha_i}$ is 
%the multiple $R^2$ obtained from regressing the shock indicator $D$ on the 
%other regressors, and $s_{\alpha_i}^2$ is... .
Propostion~\ref{prop:R1R2} states that Forecast 2 has lower forecast risk 
than Forecast 1 when  
$
  \Var(\hat{\alpha}_{\text{adj}}) 
    %- 2\E\left(\hat\eta_1 + \hat\phi_1 y_{1,T^*} + \hat\theta_1'\x_{1,T^*+1} 
    %+ \hat\beta_1'\x_{1,T^*} - \eta_1 - \phi_1 y_{1,T^*} - \theta_1'\x_{1,T^*+1}
    %- \beta_1'\x_{1,T^*}\right)\mu_{\alpha} 
    < \mu_{\alpha}^2.
$
Combining this result with \eqref{reg}, we see that Forecast 2 has lower 
forecast risk than Forecast 1 when 
%$$
%  \frac{1}{n^2(T-1)}\sum_{i=2}^{n+1} \E\left\{\frac{s_{e,j}^2}{(1-R^2_i)s_{\alpha_i}^2}\right\} 
%    + \frac{\sigma_{\alpha}^2}{n} < \mu_{\alpha}^2.
%$$
$$
  \frac{1}{n^2} \sum_{i=2}^{n+1} \E\left\{\Var(\hat{\alpha}_i|\Hist)\right\} 
    + \frac{\sigma_{\alpha}^2}{n} < \mu_{\alpha}^2.
$$
Forecast 2 is preferable to Forecast 1 asymptotically in both $T$ and $n$ 
whenever $\mu_{\alpha} \neq 0$. In finite samples, Forecast 2 is preferable to 
Forecast 1 when the $\mu_{\alpha}$ is large relative to its variability and 
overall regression variability.
